---
title: "Regression modeling"
author: "Mesut Kaya and Toine Bogers"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  rmdformats::readthedown:
    toc_depth: 3  # upto three depths of headings
css: custom.css
---


# Dataset

This R Markdown notebook is companion to the ACM TORS article *Bla * and contains the data and scripts for the regression analysis of effectiveness of candidate recommender system for recruiters. If you use this data in your own analysis, please cite the original article as follows:

> Kaya, M. & Bogers, T. (2026). Analyzing the Effects of a Human-in-the-Loop Candidate
Recommendation Algorithm at Jobindex. ACM Transactions on Recommender Systems, ?(?), ?--?.


## Initialization

We start by importing the relevant R packages, defining relevant directories and creating some helper functions for printed rounded numbers.

```{r setup, include=FALSE}

# Don't print R code by default and don't print out the R messages either.
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

```

```{r init, echo=TRUE, warning=FALSE}

# Import relevant packages.
library("glue")           # Easier string formatting
library("gtsummary")      # Provides clear summary tables of regression models
library("here")           # Easy shortcut for the root RStudio project dir
library("skimr")          # Generate useful summaries of data frames and tibbles
library("tidyverse")      # All useful tidyverse packages

# Define the relevant directories.
base_dir <- here()
processed_data_dir <- glue("{base_dir}/data/processed")

# Helper functions for properly formatting/rounding numbers.
round0 <- function(number) {
  return(scales::label_comma(accuracy = NULL)(number))
}
round1 <- function(number) {
  return(scales::label_comma(accuracy = 0.1)(number))
}
round2 <- function(number) {
  return(scales::label_comma(accuracy = 0.01)(number))
}

```


## Descriptive overview

The raw regression data can be found in `data/processed` as `raw-regression-data.tsv`. Let's read this in first:

```{r data_loading, echo=TRUE}

regression_data <- read_tsv(glue("{processed_data_dir}/raw-regression-data.tsv"),
                     show_col_types = FALSE,
                     col_types = "ccccddddddddddddddddddddDcdddddddddddDdccdcccc")

# skim(regression_data)

```

The raw regression dataset was generated from Jobindex' internal databases. It contains `r round0(nrow(regression_data))` rows and  `r ncol(regression_data)` columns. Using `skimr`, we can provide an overview of its contents:

```{r data_overview, echo=TRUE}

skim(regression_data)

```



## Computing additional variables

### Recruiter expertise

In addition to the `r ncol(regression_data)` variables, we created additional variables to, e.g., represent recruiter expertise. 

```{r additional-vars-1a, echo=TRUE}

# Bin the mean recruiter PRR scores. 
# First determine what the quantiles are.
quantiles <- regression_data %>% 
  select(recruiter_id, recruiter_prr) %>% 
  distinct() %>% 
  select(recruiter_prr) %>% 
  pull() %>% 
  quantile(., c(.25, .5, .75))
quantile_25 <- quantiles[[1]]
quantile_50 <- quantiles[[2]]
quantile_75 <- quantiles[[3]]
# Now add a column with bins.
regression_data <- regression_data %>% 
  mutate(recruiter_prr_binned = ifelse(recruiter_prr < quantile_25, "low", NA),
         recruiter_prr_binned = ifelse(recruiter_prr >= quantile_75, "high", recruiter_prr_binned),
         recruiter_prr_binned = ifelse(recruiter_prr >= quantile_25 & recruiter_prr < quantile_75, "mid", recruiter_prr_binned)) %>% 
  # Turn it into a factor.
  mutate(recruiter_prr_binned = fct_relevel(recruiter_prr_binned, c("high", "mid", "low")))
```

Show the data distribution over these bins:

```{r additional-vars-1b, echo=FALSE}

tbl_summary(regression_data, include = c(recruiter_prr_binned)) %>% as_flex_table()

```



### Company size

We also create a variable to bin the company size, according to the following rules:

- Micro-enterprises: 0-9 employees
- Small firms: 10-49 employees
- Medium-sized firms: 50-249 employees
- Large firms: 250+ employees

```{r additional-vars-2a, echo=TRUE}

# Bin the company size. 
#   Micro-enterprises: 0-9 employees.
#   Small firms: 10-49 employees.
#   Medium-sized firms: 50-249 employees.
#   Large firms: 250+ employees. 
regression_data <- regression_data %>% 
  # Now assign the bins.
  mutate(job_company_size = ifelse(job_company_num_employees < 10, "micro", NA),
         job_company_size = ifelse(job_company_num_employees >= 10 & job_company_num_employees < 50, "small", job_company_size),
         job_company_size = ifelse(job_company_num_employees >= 50 & job_company_num_employees < 250, "medium", job_company_size),
         job_company_size = ifelse(job_company_num_employees >= 250 , "large", job_company_size)) %>% 
  # Turn it into a factor.
  mutate(job_company_size = fct_relevel(job_company_size, c("large", "medium", "small", "micro")))

```

Show the data distribution over these bins:

```{r additional-vars-2b, echo=FALSE}

tbl_summary(regression_data, include = c(job_company_size)) %>% as_flex_table()

```



### Industry expertise

Finally, we also added a variable representing expertise in specific industry sectors. If 5% or more of a recruiter's total jobs are in an industry sector, we consider them an expert in that sector. Each job is classified as `yes` if they are an expert in that job's industry sector and `no` otherwise.

```{r additional-vars-4a, echo=TRUE}

# Calculate the total number of jobs per recruiter.
total_job_count <- regression_data %>% 
  dplyr::select(job_id, recruiter_id) %>% 
  dplyr::count(recruiter_id) %>% 
  dplyr::rename(total = n)

# If a recruiter has a share at or above X% in a certain industry sector, mark
# them as an expert.
expert_percentage <- 5.0
industry_expertise_per_recruiter <- regression_data %>% 
  dplyr::count(recruiter_id, job_company_industry_value) %>% 
  dplyr::inner_join(., total_job_count, by = "recruiter_id") %>% 
  mutate(perc = n / total * 100) %>% 
  # If 5% or more of your total jobs is this category, you are an expert.
  mutate(recruiter_industry_expert = ifelse(perc >= expert_percentage, "yes", "no")) %>% 
  select(recruiter_id, job_company_industry_value, recruiter_industry_expert)
# Now add these labels to the original data.
regression_data <- inner_join(regression_data, industry_expertise_per_recruiter,
                       by = c("recruiter_id", "job_company_industry_value"))

```

<!-- Show the data distribution over these bins: -->

```{r additional-vars-4b, echo=FALSE}

# tbl_summary(regression_data, include = c(recruiter_industry_expert)) %>% as_flex_table()

```



## Overview of variables

<!-- Generated using https://www.tablesgenerator.com/markdown_tables -->

The raw regression data contains the following `r ncol(regression_data)` variables:

| **Variable** | **Description** | **Variable type** |
|---|---|---|
| `job_id` |        Unique identifier for each job. This dataset has one row per job ID. | Nominal |
| `recruiter_id` | Unique identifier for each   recruiter. While multiple recruiters on a single job are possible,   recruitment is not a collaborative task at Jobindex and as a result the   dataset contains only jobs with a single recruiter. | Nominal |
| `job_complexity` | The complexity of the job as   determined by an experienced recruiter upon addition to the job queue at   Jobindex. | Nominal (‘standard’   vs. ‘complex’) |
| `job_rec_engagement` | Did the recruiter engage with   the recommendations (‘check’) or not (‘skip’)? | Nominal (‘check’   vs. ‘skip’) |
| `job_num_queries` | Number of queries issued for   this job | Ratio |
| `job_num_contacted_in_recs` | Number of candidates contacted   from the list of recommendations | Ratio |
| `job_contacted_count` | Number of candidates contacted   in the entire session | Ratio |
| `job_positive_count` | Number of contacted candidates   who responded positively | Ratio |
| `job_prr` | Positive response rate (positive   responses / all contact candidates) | Ratio |
| `job_negative_count` | Number of contacted candidates   who responded negatively | Ratio |
| `job_nrr` | Negative response rate (negative   responses / all contact candidates) | Ratio |
| `job_all_cv_views` | Number of CVs viewed in the   entire session | Ratio |
| `job_rec_cv_views` | Number of CVs viewed from the   recommendation list | Ratio |
| `job_non_rec_cv_views` | Number of CVs viewed during the   manual searches | Ratio |
| `job_rec_cv_adds` | Number of CVs   shortlisted/contacted from the recommendation list | Ratio |
| `job_non_rec_cv_adds` | Number of CVs   shortlisted/contacted in the entire session | Ratio |
| `job_all_cv_adds` | Number of CVs   shortlisted/contacted during the manual searches | Ratio |
| `job_rec_cv_clicks` | Number of CVs clicked from the   recommendation list | Ratio |
| `job_all_cv_clicks` | Number of CVs clicked in the   entire session | Ratio |
| `job_non_rec_cv_clicks` | Number of CVs clicked during the   manual searches | Ratio |
| `job_contacted_in_recs` | ? | Ratio |
| `job_contacted_from_recs` | ? | Ratio |
| `job_pos_from_recs` | ? | Ratio |
| `job_pos_in_recs` | ? | Ratio |
| `job_month` | The month in which the job was   processed (recorded as the first day of that month) | Date |
| `job_ui_change` | Was this job published before   ('Before') or after ('After') the UI change? | Nominal (‘Before’   vs. ‘After’) |


<!-- #  * job_complexity                How complex is the job? ('complex' vs. 'standard') -->
<!-- #  * job_rec_engagement            Engagement with recommendations? ('view' vs. 'skip') -->
<!-- #  * job_ui_change                 Was this before/after the UI change? ('Before' vs. 'After') -->
<!-- # -->
<!-- #  * job_num_queries               How actively did they search? -->
<!-- #  * job_all_cv_views              How many CVs did they view? -->
<!-- #  * job_contacted_count           How many did they contact? -->
<!-- #  * job_company_industry_value    Industry sector of the job -->
<!-- #  * job_company_name              Company of the job (potential random effect?) -->
<!-- # -->
<!-- #  * recruiter_view_percentage     Mean % of jobs this recruiter views the recs for -->
<!-- #  * recruiter_prr                 Mean PRR for this recruiter -->
<!-- #  * recruiter_prr_binned          Binned version of 'recruiter_prr' (fixed no. of levels, so fixed effect) -->
<!-- #  * recruiter_months_experience   No. of months this recruiter has been working (until end of dataset) -->
<!-- #  * recruiter_exp_binned          Binned version of 'recruiter_months_experience' (fixed no. of levels, so fixed effect) -->

# Regression analysis

## Mixed-effects model

In order to determine which factor(s) best predict the positive response rate, we constructed a two-level linear model (or mixed-effects model) with completed jobs (Level 1) nested within recruiters (Level 2). The outcome variable was the job-specific PRR score. The expectation was that a meaningful part of the differences in PRR scores could be explained by the differences between recruiters. We carefully added potential predictors in line with our research questions as both fixed and random effects only if they resulted in a significant decrease in BIC score. However, the Intra-Class Correlation (ICC) of the best-performing model was 0.020, which suggests that only 2% of the variance can explained by our grouping structure. This suggests that a mixed-effects model is not necessary due to the lack of variance at the higher level.



## Multi-variate regression model

We therefore instead estimated a one-level multiple-regression model, again by carefully adding potential predictors in line with our research questions only if they resulted in a significant decrease in BIC score.

```{r, echo=TRUE}

linear_model_final <- lm(job_prr ~ job_rec_engagement + job_complexity + job_ui_change + 
                     recruiter_prr + job_complexity:job_ui_change +  
                     recruiter_industry_expert + job_all_cv_views, 
                   data = regression_data, na.action = na.exclude)
```

The coefficients of this regression model are shown below:

```{r, echo=FALSE}
tbl_regression(linear_model_final,
               estimate_fun = label_style_number(digits = 3),    # for estimates/CIs
               pvalue_fun = label_style_pvalue(digits = 3)) %>%  # for p-values) 
   as_flex_table()

```

The regression equation was significant ($F$(7, 36029) = 326.3, $p$ < .001) with the model explaining 6\% of the variance in PRR scores ($R^2$ = .059). 




The resulting multiple-regression model predicts PRR scores from the following predictors:

- Engagement with the recommendations (`job_rec_engagement`) was a significant predictor ($\beta$ = 0.0123, $p$ < .001) as checking the recommendations significantly increases PRR for a job.
- Job complexity (`job_complexity`) was a significant predictor ($\beta$ = -0.0079, $p$ < .001) as standard jobs have significantly lower PRR than complex jobs.
- The mean PRR score achieved by the recruiter (`recruiter_prr`) in question was a significant predictor ($\beta$ = 1.0011, $p$ < .001). This variable is a proxy for recruiter expertise and higher mean PRR scores translate into higher PRR scores for individual jobs.    
- Recruiters typically have a set of industry sectors that they are specialized in. We define all sectors that make up 5% or more of a recruiter's work as one of their specializations and flagged this for each job as a binary variable. This predictor (`recruiter_industry_expert`) was a significant predictor ($\beta$ = -0.0047, $p$ < .001). Perhaps counterintuitively, jobs from industry sectors that a recruiter specialized in resulted in lower PRR scores. 
- The number of candidate CVs that the recruiter viewed for a job (`job_all_cv_views`) was a significant predictor ($\beta$ = $2.1 \cdot 10^{-5}$, $p$ < .001). Viewing more CVs results in higher PRR scores although the effect is very small.
- Whether a job was completed before or after the UI change (`job_ui_change`) was not a significant predictor ($\beta$ = 0.0006, $p$ < .249), even though removing this predictor from the model reduced model fit significantly. 
- The interaction between job complexity (`job_complexity`) and the timing of the UI change (`job_ui_change`) was a significant predictor ($\beta$ = -0.0189, $p$ < .001). In particular, it shows that standard jobs achieve significantly lower PRR rates before the UI change than the other combinations. We believe this is because after the UI change there was a bigger share of standard jobs  being checked, resulting in a bigger overall PRR increase for standard jobs compared to before.


| b | a |
|---|---|
| job_complexity              | How complex is the job? ('complex' vs. 'standard') |
| job_rec_engagement          | Engagement with recommendations? ('view' vs. 'skip') |
| job_ui_change               | Was this before/after the UI change? ('Before' vs. 'After') |
| job_num_queries             | How actively did they search? |
| job_all_cv_views            | How many CVs did they view? |
| job_contacted_count         | How many did they contact? |
| job_company_industry_value  | Industry sector of the job |
| job_company_name            | Company of the job (potential random effect?) |
| recruiter_view_percentage   | Mean % of jobs this recruiter views the recs for |
| recruiter_prr               | Mean PRR for this recruiter |
| recruiter_prr_binned        | Binned version of 'recruiter_prr' (fixed no. of levels, so fixed effect) |
| recruiter_months_experience | No. of months this recruiter has been working (until end of dataset) |
| recruiter_exp_binned        | Binned version of 'recruiter_months_experience' (fixed no. of levels, so fixed effect) |

