---
title: "Regression modeling"
author: "Mesut Kaya and Toine Bogers"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  rmdformats::readthedown:
    toc_depth: 3  # upto three depths of headings
css: custom.css
---


# Dataset

## Initialization

We start by importing the relevant R packages, defining relevant directories and creating some helper functions for printed rounded numbers.

```{r setup, include=FALSE}

# Don't print R code by default and don't print out the R messages either.
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

```

```{r init, echo=TRUE}

# Import relevant packages.
library("glue")           # Easier string formatting
library("gtsummary")
library("broom")
library("broom.helpers")
library("here")           # Easy shortcut for the root RStudio project dir
library("skimr")          # Generate useful summaries of data frames and tibbles.
library("tidyverse")      # All useful tidyverse packages

# Define the relevant directories.
base_dir <- here()
processed_data_dir <- glue("{base_dir}/data/processed")

# Helper functions for properly formatting/rounding numbers.
round0 <- function(number) {
  return(scales::label_comma(accuracy = NULL)(number))
}
round1 <- function(number) {
  return(scales::label_comma(accuracy = 0.1)(number))
}
round2 <- function(number) {
  return(scales::label_comma(accuracy = 0.01)(number))
}

```


## Descriptive overview

The raw regression data can be found in `data/processed` as `raw-regression-data.tsv`. Let's read this in first:

```{r data_loading, echo=TRUE}

raw_data <- read_tsv(glue("{processed_data_dir}/raw-regression-data.tsv"),
                     show_col_types = FALSE,
                     col_types = "ccccddddddddddddddddddddDcdddddddddddDdccdcccc")

# skim(raw_data)

```

The raw regression dataset was generated from Jobindex' internal databases. It contains `r round0(nrow(raw_data))` rows and  `r ncol(raw_data)` columns. Using `skimr`, we can provide an overview of its contents:

```{r data_overview, echo=TRUE}

skim(raw_data)

```



## Computing additional variables

### Recruiter expertise

In addition to the `r ncol(raw_data)` variables, we created additional variables to, e.g., represent recruiter expertise. 

```{r additional-vars-1a, echo=TRUE}

# Bin the mean recruiter PRR scores. 
# First determine what the quantiles are.
quantiles <- raw_data %>% 
  select(recruiter_id, recruiter_prr) %>% 
  distinct() %>% 
  select(recruiter_prr) %>% 
  pull() %>% 
  quantile(., c(.25, .5, .75))
quantile_25 <- quantiles[[1]]
quantile_50 <- quantiles[[2]]
quantile_75 <- quantiles[[3]]
# Now add a column with bins.
raw_data <- raw_data %>% 
  mutate(recruiter_prr_binned = ifelse(recruiter_prr < quantile_25, "low", NA),
         recruiter_prr_binned = ifelse(recruiter_prr >= quantile_75, "high", recruiter_prr_binned),
         recruiter_prr_binned = ifelse(recruiter_prr >= quantile_25 & recruiter_prr < quantile_75, "mid", recruiter_prr_binned)) %>% 
  # Turn it into a factor.
  mutate(recruiter_prr_binned = fct_relevel(recruiter_prr_binned, c("high", "mid", "low")))
```

Show the data distribution over these bins:

```{r additional-vars-1b, echo=FALSE}

tbl_summary(raw_data, include = c(recruiter_prr_binned)) %>% as_flex_table()

```



### Company size

We also create a variable to bin the company size, according to the following rules:

- Micro-enterprises: 0-9 employees
- Small firms: 10-49 employees
- Medium-sized firms: 50-249 employees
- Large firms: 250+ employees

```{r additional-vars-2a, echo=TRUE}

# Bin the company size. 
#   Micro-enterprises: 0-9 employees.
#   Small firms: 10-49 employees.
#   Medium-sized firms: 50-249 employees.
#   Large firms: 250+ employees. 
raw_data <- raw_data %>% 
  # Now assign the bins.
  mutate(job_company_size = ifelse(job_company_num_employees < 10, "micro", NA),
         job_company_size = ifelse(job_company_num_employees >= 10 & job_company_num_employees < 50, "small", job_company_size),
         job_company_size = ifelse(job_company_num_employees >= 50 & job_company_num_employees < 250, "medium", job_company_size),
         job_company_size = ifelse(job_company_num_employees >= 250 , "large", job_company_size)) %>% 
  # Turn it into a factor.
  mutate(job_company_size = fct_relevel(job_company_size, c("large", "medium", "small", "micro")))

```

Show the data distribution over these bins:

```{r additional-vars-2b, echo=FALSE}

tbl_summary(raw_data, include = c(job_company_size)) %>% as_flex_table()

```



### Industry expertise

Finally, we also added a variable representing expertise in specific industry sectors. If 5% or more of a recruiter's total jobs are in an industry sector, we consider them an expert in that sector. Each job is classified as `yes` if they are an expert in that job's industry sector and `no` otherwise.

```{r additional-vars-4a, echo=TRUE}

# Calculate the total number of jobs per recruiter.
total_job_count <- raw_data %>% 
  dplyr::select(job_id, recruiter_id) %>% 
  dplyr::count(recruiter_id) %>% 
  dplyr::rename(total = n)

# If a recruiter has a share at or above X% in a certain industry sector, mark
# them as an expert.
expert_percentage <- 5.0
industry_expertise_per_recruiter <- raw_data %>% 
  dplyr::count(recruiter_id, job_company_industry_value) %>% 
  dplyr::inner_join(., total_job_count, by = "recruiter_id") %>% 
  mutate(perc = n / total * 100) %>% 
  # If 5% or more of your total jobs is this category, you are an expert.
  mutate(recruiter_industry_expert = ifelse(perc >= expert_percentage, "yes", "no")) %>% 
  select(recruiter_id, job_company_industry_value, recruiter_industry_expert)
# Now add these labels to the original data.
raw_data <- inner_join(raw_data, industry_expertise_per_recruiter,
                       by = c("recruiter_id", "job_company_industry_value"))

```

<!-- Show the data distribution over these bins: -->

```{r additional-vars-4b, echo=FALSE}

# tbl_summary(raw_data, include = c(recruiter_industry_expert)) %>% as_flex_table()

```



## Overview of variables

<!-- Generated using https://www.tablesgenerator.com/markdown_tables -->

The raw regression data contains the following `r ncol(raw_data)` variables:

| **Variable** | **Description** | **Variable type** |
|---|---|---|
| `job_id` |        Unique identifier for each job. This dataset has one row per job ID. | Nominal |
| `recruiter_id` | Unique identifier for each   recruiter. While multiple recruiters on a single job are possible,   recruitment is not a collaborative task at Jobindex and as a result the   dataset contains only jobs with a single recruiter. | Nominal |
| `job_complexity` | The complexity of the job as   determined by an experienced recruiter upon addition to the job queue at   Jobindex. | Nominal (‘standard’   vs. ‘complex’) |
| `job_rec_engagement` | Did the recruiter engage with   the recommendations (‘check’) or not (‘skip’)? | Nominal (‘check’   vs. ‘skip’) |
| `job_num_queries` | Number of queries issued for   this job | Ratio |
| `job_num_contacted_in_recs` | Number of candidates contacted   from the list of recommendations | Ratio |
| `job_contacted_count` | Number of candidates contacted   in the entire session | Ratio |
| `job_positive_count` | Number of contacted candidates   who responded positively | Ratio |
| `job_prr` | Positive response rate (positive   responses / all contact candidates) | Ratio |
| `job_negative_count` | Number of contacted candidates   who responded negatively | Ratio |
| `job_nrr` | Negative response rate (negative   responses / all contact candidates) | Ratio |
| `job_all_cv_views` | Number of CVs viewed in the   entire session | Ratio |
| `job_rec_cv_views` | Number of CVs viewed from the   recommendation list | Ratio |
| `job_non_rec_cv_views` | Number of CVs viewed during the   manual searches | Ratio |
| `job_rec_cv_adds` | Number of CVs   shortlisted/contacted from the recommendation list | Ratio |
| `job_non_rec_cv_adds` | Number of CVs   shortlisted/contacted in the entire session | Ratio |
| `job_all_cv_adds` | Number of CVs   shortlisted/contacted during the manual searches | Ratio |
| `job_rec_cv_clicks` | Number of CVs clicked from the   recommendation list | Ratio |
| `job_all_cv_clicks` | Number of CVs clicked in the   entire session | Ratio |
| `job_non_rec_cv_clicks` | Number of CVs clicked during the   manual searches | Ratio |
| `job_contacted_in_recs` | ? | Ratio |
| `job_contacted_from_recs` | ? | Ratio |
| `job_pos_from_recs` | ? | Ratio |
| `job_pos_in_recs` | ? | Ratio |
| `job_month` | The month in which the job was   processed (recorded as the first day of that month) | Date |
| `job_ui_change` | Was this job published before   ('Before') or after ('After') the UI change? | Nominal (‘Before’   vs. ‘After’) |


<!-- #  * job_complexity                How complex is the job? ('complex' vs. 'standard') -->
<!-- #  * job_rec_engagement            Engagement with recommendations? ('view' vs. 'skip') -->
<!-- #  * job_ui_change                 Was this before/after the UI change? ('Before' vs. 'After') -->
<!-- # -->
<!-- #  * job_num_queries               How actively did they search? -->
<!-- #  * job_all_cv_views              How many CVs did they view? -->
<!-- #  * job_contacted_count           How many did they contact? -->
<!-- #  * job_company_industry_value    Industry sector of the job -->
<!-- #  * job_company_name              Company of the job (potential random effect?) -->
<!-- # -->
<!-- #  * recruiter_view_percentage     Mean % of jobs this recruiter views the recs for -->
<!-- #  * recruiter_prr                 Mean PRR for this recruiter -->
<!-- #  * recruiter_prr_binned          Binned version of 'recruiter_prr' (fixed no. of levels, so fixed effect) -->
<!-- #  * recruiter_months_experience   No. of months this recruiter has been working (until end of dataset) -->
<!-- #  * recruiter_exp_binned          Binned version of 'recruiter_months_experience' (fixed no. of levels, so fixed effect) -->

# Regression analysis

## Multi-variate regression model

Similar to the mixed-effects model, we built up the linear model by checking for significantly lower BIC scores. The final model was generated as follows:

```{r, echo=TRUE}

linear_model_final <- lm(job_prr ~ job_rec_engagement + job_complexity + job_ui_change + 
                     recruiter_prr + job_complexity:job_ui_change +  
                     recruiter_industry_expert + job_all_cv_views, 
                   data = raw_data, na.action = na.exclude)
```


```{r, echo=FALSE, warning=FALSE}

# summary(linear_model_final)
# broom::tidy(linear_model_final)
# performance::model_performance(linear_model_final)

BIC_linear_model_final <- performance::model_performance(linear_model_final) %>% 
  broom::tidy() %>% 
  dplyr::filter(column == "BIC") %>% 
  dplyr::select(mean) %>% 
  dplyr::pull()

# Feed model to gt_summary

# bind_cols(
#   tidy(lm_model_final) %>%
#     mutate(estimate = sprintf("%.4f", estimate),
#            std.error = sprintf("%.4f", std.error),
#            statistic = sprintf("%.3f", statistic),
#            p.value = sprintf("%.3f", p.value)),
#   confint(lm_model_final, level=0.95) %>% as_tibble() %>% dplyr::rename(low = `2.5 %`, high = `97.5 %`) %>% 
#     mutate(confint = glue("[{sprintf('%.4f', low)}, {sprintf('%.4f', high)}]"))
# ) %>% 
#   mutate(latex_string = glue("&   {estimate}  & {std.error}  & {statistic} & {p.value}  & {confint} \\")) %>% 
#   select(1, 9)

```



This model had a BIC value of `r round2(BIC_linear_model_final)`, which was significanty higher than ...


```{r, echo=FALSE}
tbl_regression(linear_model_final) %>% as_flex_table()

```




## Mixed-effects model

A two-level linear model was estimated with job-specific PRR scores (Level 1, `job_prr`) nested within recruiters (Level 2, `recruiter_id`). The outcome variable was the job-specific PRR score, predicted by treatment group and time, with random intercepts for individuals and random slopes for time.

We considered the following Level 1 variables:

| b | a |
|---|---|
| job_complexity              | How complex is the job? ('complex' vs. 'standard') |
| job_rec_engagement          | Engagement with recommendations? ('view' vs. 'skip') |
| job_ui_change               | Was this before/after the UI change? ('Before' vs. 'After') |
| job_num_queries             | How actively did they search? |
| job_all_cv_views            | How many CVs did they view? |
| job_contacted_count         | How many did they contact? |
| job_company_industry_value  | Industry sector of the job |
| job_company_name            | Company of the job (potential random effect?) |
| recruiter_view_percentage   | Mean % of jobs this recruiter views the recs for |
| recruiter_prr               | Mean PRR for this recruiter |
| recruiter_prr_binned        | Binned version of 'recruiter_prr' (fixed no. of levels, so fixed effect) |
| recruiter_months_experience | No. of months this recruiter has been working (until end of dataset) |
| recruiter_exp_binned        | Binned version of 'recruiter_months_experience' (fixed no. of levels, so fixed effect) |

We constructed the 

Explain that we build the model iteratively starting with x and x, adding predictors if they increased the BIC score significantly according to an ANOVA comparison. If predictors became insignificant later on, we removed them if the model fit improved. We tried different interaction effects, but only one 'worked'. We added all fixed effects as random effects, but only in three cases did this improve the model fit. The following predictors did not improve model fit at all: 

